Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-10-15 Background.tex
edited second half

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{(BladeLength - Rx_y)}\right)
	\label{eq:theory_90_limited}
\end{equation}

@@

Figure \ref{fig:3D_model_270az} depicts the transmitter at a 270\textdegree \space azimuth angle, which places it in-line with the position of the receiver and the current location of the rotating blades. This placement of the transmitter behind the helicopter is different from the situation in Figure \ref{fig:3D_model_90az} because the value of $\theta_r$ is now not limited by the tip of the rotor and can go to smaller elevation angles.

@@

%formulate the geometric equations
The equation for modeling the scenario for a transmitter located at 270\textdegree \space azimuth is similar to the equations for calculating the 90\textdegree \space scenario, the limiting factor is now the reflection radius, because as the transmitter moves farther away the elevation angle is increasing but the reflection radius gets smaller.

>>>>

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{([0,\dots, (BladeLength - Rx_y)])}\right)
	\label{eq:theory_90_limited}
\end{equation}

@@

Figure \ref{fig:3D_model_270az} depicts the transmitter at a 270\textdegree \space azimuth angle, which places it in-line with the position of the receiver and the current location of the rotating blades. This placement of the transmitter behind the helicopter is different from the situation in Figure \ref{fig:3D_model_90az} because the value of $\theta_r$ is now limited by the  rotor hub and maximum reflection radius.

@@

%formulate the geometric equations
The equation for modeling the scenario for a transmitter located at 270\textdegree \space azimuth is similar to the equations for calculating the 90\textdegree \space scenario where $f_{DopplerMax}$ is now calculated with a smaller reflection radius of 2m and $\theta_{Limited}$ is now calculated by

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{([0, \dots, BladeLength)}\right)
	\label{eq:theory_270_limited}
\end{equation}

### 2016-10-15 Conclusion.tex
edited second half

\section{Future Work}
%Real world experiment needs to be conducted starting with small scale and moving towards a large scale implementation on a helicopter platform.
The first thing that needs to be done in continuing this research is to perform a real world experiment to officially evaluate the ray tracing application. To do this appropriate hardware needs to be defined in the form of a software defined radio for both the receiver and transmitter along with a software workflow such as GNU Radio. Next a small scale environment would need to be set up, including the above hardware along with some form of rotating metal blade or thin plate rotating at a sufficient rate to produce a doppler spread similar to those defined in the simulation. After that data is collected and analyzed a full scale experiment would need to be conducted including a real helicopter platform and an environment with large expanses of flat ground.

@@

%Improve computation speed by employing multithreading or computation on GPU Architecture for reduced simulation time, not necessary but would be more convenient.
Another improvement on the simulation would be to increase the speed of computation. This could be done by implementing multi-threading or, since raytracing is heavily parallelizable, a GPU based architecture. The multi-threading option would require minor code modifications but realizing the simulation on a GPU would require a full rewrite of the software to take advantage of the GPU's abilities.

@@

%Because there is more information in the received envelopes besides the maximum and minimum the specific shape could be used to better estimate azimuth and elevation without the need to complete a full revolution as described in the proposed method.
The estimation methods described in this thesis rely only on the maximum and minimum values of the doppler envelopes calculated but their shape could also be of particular interest as well. When varying the parameters of azimuth and elevation the shape of the envelope changes in a distinctive manner, and by using a training algorithm the position of the transmitter might be able to be tracked in time by evaluating the change in shape. where currently if we want to locate the transmitter in a successive fashion a full revolution of the helicopter platform is needed to resolve position ambiguity.

>>>>

\section{Future Work}
%Real world experiment needs to be conducted starting with small scale and moving towards a large scale implementation on a helicopter platform.
The first thing that needs to be done in continuing this research is to perform a real world experiment to officially evaluate the ray tracing application. To do this appropriate hardware needs to be defined in the form of a software defined radio for both the receiver and transmitter along with a software workflow such as GNU Radio. Next a small scale environment would need to be set up, with some form of rotating metal blade or thin plate rotating at a sufficient rate to produce a doppler spread similar to those defined in the simulation. After that data is collected and analyzed a full scale experiment would need to be conducted including a real helicopter platform and an environment with large expanses of flat ground.

@@

%Improve computation speed by employing multithreading or computation on GPU Architecture for reduced simulation time, not necessary but would be more convenient.
Another improvement on the simulation would be to increase the speed of computation. This could be done by implementing multi-threading or, since raytracing is heavily parallelizable, a Graphics Processing Unit (GPU) based architecture. The multi-threading option would require minor code modifications but realizing the simulation on a GPU would require a full rewrite of the software to take advantage of the GPU's abilities.

@@

%Because there is more information in the received envelopes besides the maximum and minimum the specific shape could be used to better estimate azimuth and elevation without the need to complete a full revolution as described in the proposed method.
The estimation methods described in this thesis rely only on the maximum and minimum values of the doppler envelopes calculated but their shape could also be of particular interest as well. When varying the parameters of azimuth and elevation the shape of the envelope changes in a distinctive manner, by using a training algorithm the position of the transmitter might be able to be tracked in time by evaluating the change in shape. Under the defined estimation methods a full revolution of the helicopter platform is needed to resolve position ambiguity.

### 2016-10-15 Radio_Propagation_Simulation.tex
edited second half

\subsection{Receiver}
%physical receiver dimensions and antenna chars
The receiver object is modeled as a point in 3D space with a surrounding boundary layer in the shape of a sphere. This is because the receiver is assumed to have an omnidirectional antenna which will receive radio waves from all directions uniformly. The receiver will record any ray that hits the boundary layer within each frame of the simulation to create a sample of data at that instance in time.
%processing in each frame
The sample is produced using superposition of all the rays that are acting on the receiver at that instance in time which are then extrapolated to produce a clear result.
%realism
The receiver is configurable allowing the user to set a desired center frequency along with the sampling rate of the simulated hardware. The sampling rate will determine several simulation parameters in the ray tracer due to the need to associate this rate with the change in the rotor position for the next frame. The consequence of this is that increasing the sample rate will produce higher resolution results but will increase the computational overhead of the simulation.

>>>>

\subsection{Receiver}
%physical receiver dimensions and antenna chars
The receiver object is modeled as a point in 3D space with a surrounding boundary layer in the shape of a sphere. This is because the receiver is assumed to have an omnidirectional antenna which will receive radio waves from all directions uniformly. The receiver will record any ray that hits the boundary layer within each frame of the simulation to create a sample of data at that instance in time.
%processing in each frame
The sample is produced using superposition of all the rays that are acting on the receiver at that instance in time which are then extrapolated to produce a clear result.
%realism
The receiver is configurable allowing the user to set a desired center frequency along with the sampling rate of the simulated hardware. The sampling rate will determine several simulation parameters in the ray tracer due to the need to associate this rate with the change in the rotor position for the next frame. The consequence of this is that increasing the sample rate will produce higher resolution results but, will increase the computational overhead of the simulation.

### 2016-10-15 Ray_Tracing_Software.tex
edited second half

The output file also needs to be defined with a unique name that changes, this is important because the program will just write over the old file if you do not have a unique naming scheme. this could be just an increasing number or related to one of the variable sweeps.

@@

this is all provided in a single string of comma separated values called a .CSV file. MATLAB has a built in function to read CSV data.

@@

after all the data is read into MATLAB the data can be split into two vectors, one pertaining to the physical simulation parameters and the other to the received signal.

>>>>

The output file also needs to be defined with a unique name that changes, this is important because the program will just write over the old file if you do not have a unique naming scheme. This could be just an increasing number or related to one of the variable sweeps.

@@

This data is all provided in a single string of comma separated values called a .CSV file. MATLAB has a built in function to read CSV data.

@@

After all the data is read into MATLAB the data can be split into two vectors, one pertaining to the physical simulation parameters and the other to the received signal.

### 2016-10-15 Results_new.tex
edited second half

\section{Introduction}
%introduce the results section and what processing will be done
The information gathered from the previous section is used to estimate the azimuth and elevation angles of the transmitter. First the azimuth angle will be estimated which will provide the information necessary for calculating the elevation angle. The results of the estimation techniques are then analyzed to determine their accuracy compared to the actual.

@@

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which a Doppler envelope and subsequent maximum and minimum values are calculated. After a complete rotation the Doppler maximum and minimum values produce a sinusoidal pattern as shown in Figure \ref{fig:tx_azimuth_angle_200} without the assumption of knowing the starting azimuth angle. With this data we need to find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum envelope data is superimposed onto the maximum envelope data as shown in Figure \ref{fig:azimuth_estimation_max_vs_absMin}.

@@

From Figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. By computing the difference

@@

While Figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself, therefore by computing the correct Doppler using the $Max Upper Envelope$ and the $| Min Lower Envelope |$. This is done by evaluating the difference at each angle and if the difference is negative the $| Min Lower Envelope |$ is used and if positive the $Max Upper Envelope$.

@@

The correct Doppler is shown in Figure \ref{fig:correct_fd} which displays the amount of Doppler that is currently acting on the signal for that specific azimuth angle. The data gathered contains a global minimum at 90\textdegree \space and a second local minimum at 270\textdegree. The 90\textdegree \space can be found easily by taking the minimum of the correct Doppler signal

@@

But in order to find the local minimum at 270\textdegree \space we need to use a different method. To do this we can use the peak detection function within MATLAB to locate both the 90\textdegree \space and 270\textdegree \space point. By taking the inverse of the correct Doppler and tuning the peak detection function with the maximum and minimum values to determine accurate prominence.

@@

Figure \ref{fig:ave_power}  shows the average power at each angle sample with the angles 90\textdegree \space and 270\textdegree \space denoted as red x's. From the plot we can see that the average power at 90\textdegree \space is much smaller than the the power at 270\textdegree, so this inequality will be used to resolve ambiguity in the azimuth angle.

@@

From Table \ref{tab:az_error_and_percent} we can see that there is no error in the estimation of the 90\textdegree \space point over a variety of Altitude and elevation angles that were tested. But because of the algorithm could potentially pick the 270\textdegree \space we will analyze the difference in peak heights and the power difference to evaluate the estimation further.

@@

From Table \ref{tab:peaks_and_power} we can see that at the larger elevation angles the peak percent difference are significant to determine which is correctly the 90\textdegree \space value, but from the Peak frequency difference at the large elevation angles the difference is in the tens of Hz so in practical applications it is possible to confuse the estimation by just taking the minimum frequency so other values need to play a role such as the power.

@@

\section{Transmitter Elevation Angle Estimation} \label{sec:transmitter_elevation_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
Using the information gathered from the azimuth estimation procedure we can compute the elevation angle of the transmitter. In the process of gathering azimuth information the maximum and minimum envelopes follow a sinusoidal pattern that contains a maximum Doppler information at approximately 180\textdegree \space and 0\textdegree \space azimuth angles. Using that maximum value along with the correct Doppler and the assumption that the maximum will occur near the tip of the rotor blade we can approximate the respected reflection radii off the blade as the azimuth angle changes. This is done through 
%reflection radius aproximation equation
\begin{equation}
	R_{Estimate} = BladeLength \left(\frac{f_{DopplerCorrect}(\theta_{Az})}{Max(f_{DopplerCorrect})}\right)
	\label{eqn:r_estimate}
\end{equation}

@@

Figure \ref{fig:elevation_comparason_azimuth_range} shows the estimation of the elevation angle $\theta_{El}$ with respect to the azimuth angle, what it shows is that the elevation angle remains constant with respect to azimuth angle. which means that the estimation in the reflection radius is correctly compensating for the azimuth angle. The percent error off the actual elevation angle is shown in Figure \ref{fig:percent_error_elevation_azimuth_range} which is very small and consistent over the azimuth range.

@@

From Table \ref{tab:elevation_error_percent} elevation estimation error is extremely small for the given actual elevation angles. But there is a trend that is evident in the data that as the elevation angle decreases the amount of error is increasing.

@@

This error is caused by a limitation on the estimation is that the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is calculated by

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree \space. The maximum and minimum envelope data is shown in Figure \ref{fig:envelope_180deg}

@@

Using the estimation method described above, Figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till about 18\textdegree \space or 20\textdegree \space where it becomes noisy. This is due to the inability of the envelope estimation to resolve the correct maximum Doppler for that range. By going back to Figure \ref{fig:tx_range_elevation_rel} we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

Figure \ref{fig:percent_error_elevation_180deg} shows the percent error with respect to the elevation angle and it starts to become unusable around 18\textdegree \space or 20\textdegree \space where the percent error starts to shoot past acceptable ranges.

@@

Since $\theta_{Az}$ is unknown during the calculation of this equation the correct starting phase can be determined by matching it to the $EnvelopeDifference$ with a priori information that the cosign should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to Figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to the unpitched results.

>>>>

\section{Introduction}
%introduce the results section and what processing will be done
The information gathered from the previous section is used to estimate the azimuth and elevation angles of the transmitter. First the azimuth angle will be estimated which will provide the information necessary for calculating the elevation angle. The results of the estimation techniques are then analyzed to determine their accuracy.

@@

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which a Doppler envelope and subsequent maximum and minimum values are calculated. After a complete rotation the Doppler maximum and minimum values produce a sinusoidal pattern as shown in Figure \ref{fig:tx_azimuth_angle_200} without knowing the starting azimuth angle. With this data we find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum envelope data is superimposed onto the maximum envelope data as shown in Figure \ref{fig:azimuth_estimation_max_vs_absMin}.

@@

From Figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. By computing the difference of the two

@@

While Figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself. Therefore by finding the correct Doppler for each azimuth, from the $Max Upper Envelope$ and the $| Min Lower Envelope |$, we can determine an unambiguous reference. This is done by evaluating the difference at each angle, and if the difference is negative the $| Min Lower Envelope |$ is used and if positive the $Max Upper Envelope$.

@@

The correct Doppler is shown in Figure \ref{fig:correct_fd} which displays the amount of Doppler that is currently acting on the signal for that specific azimuth angle. The data gathered contains a global minimum at 90\textdegree \space and a second local minimum at 270\textdegree. The 90\textdegree \space can be found easily by taking the minimum of the correct Doppler signal.

@@

In order to find the local minimum at 270\textdegree \space we need to use a different method.  We can do this through the peak detection function within MATLAB to locate both the 90\textdegree \space and 270\textdegree \space point. By taking the inverse of the correct Doppler and tuning the peak detection function with the maximum and minimum values to determine accurate prominence.

@@

Figure \ref{fig:ave_power}  shows the average power at each angle sample with the angles 90\textdegree \space and 270\textdegree \space denoted as red x's. From the plot we can see that the average power at 90\textdegree \space is much smaller than at 270\textdegree, so this inequality will be used to resolve ambiguity in the azimuth angle.

@@

From Table \ref{tab:az_error_and_percent} we can see that there is no error in the estimation of the 90\textdegree \space point over a variety of altitude and elevation angles that were tested. But because the algorithm could potentially pick the 270\textdegree \space we will analyze the difference in peak heights and the power difference to evaluate the estimation further.

@@

From Table \ref{tab:peaks_and_power} we can see that at the larger elevation angles the peak percent difference are significant to determine which is correctly the 90\textdegree \space value. Although from the frequency difference, at the large elevation angles, it is possible to confuse the estimation by just taking the minimum frequency. Therefore other values need to play a role such as the received power.

@@

\section{Transmitter Elevation Angle Estimation} \label{sec:transmitter_elevation_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
Using the information gathered from the azimuth estimation procedure, we can compute the elevation angle of the transmitter. In the process of gathering azimuth information, the maximum and minimum envelopes follow a sinusoidal pattern that contains a maximum Doppler information at approximately 180\textdegree \space and 0\textdegree \space azimuth angles. Using that maximum value, along with the correct Doppler and the assumption that the maximum will occur near the tip of the rotor blade, we can approximate the respected reflection radii off the blade as the azimuth angle changes. This is done through 
%reflection radius aproximation equation
\begin{equation}
	R_{Estimate} = BladeLength \left(\frac{f_{DopplerCorrect}(\theta_{Az})}{Max(f_{DopplerCorrect})}\right)
	\label{eqn:r_estimate}
\end{equation}

@@

Figure \ref{fig:elevation_comparason_azimuth_range} shows the estimation of the elevation angle $\theta_{El}$ with respect to the azimuth angle. What it shows is that the elevation angle remains constant with respect to azimuth angle. Which means that the estimation in the reflection radius is correctly compensating for the azimuth angle. The percent error off the actual elevation angle is shown in Figure \ref{fig:percent_error_elevation_azimuth_range} which is very small but consistent over the azimuth range.

@@

From Table \ref{tab:elevation_error_percent} we can see elevation estimation error is extremely small for the given elevation angles. There is a trend that is evident in the data that as the elevation angle decreases the amount of error is increasing.

@@

The error in elevation is caused by a limitation on the estimation. Where the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is calculated by

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree. The maximum and minimum envelope data is shown in Figure \ref{fig:envelope_180deg}

@@

Using the estimation method described above, Figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till elevation angles of 18\textdegree \space or 20\textdegree, where it becomes noisy. This is due to the inability of the envelope estimation to resolve the correct maximum Doppler for that range. By going back to Figure \ref{fig:tx_range_elevation_rel} we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

Figure \ref{fig:percent_error_elevation_180deg} shows the percent error with respect to the elevation angle. The estimation starts to become unusable around 18\textdegree \space or 20\textdegree \space where the percent error starts to shoot past acceptable ranges.

@@

Since $\theta_{Az}$ is unknown during the calculation of equation \ref{eqn:fd_pitch_correct} the correct starting phase can be determined by matching it to the $EnvelopeDifference$ and that the cosine should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to Figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to the unpitched results.

### 2016-10-15 Simulation.tex
edited second half

The rotor blade length from Figure \ref{fig:rx_position_image} is 7.5m so the receiver will be varied from 0m to 9m in 1m increments. The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of the position of the receiver.

@@

Figure \ref{fig:rx_position} shows maximum Doppler vs the position of the receiver. From this graph the maximum Doppler frequency changes significantly as the receiver is moved from directly under neath the rotor to past the tip of the rotor at 9m.The position of the receiver also effects the resulting signal envelope shapes shown in Figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

@@

Comparing Figure \ref{fig:rx_position_0m} with \ref{fig:rx_position_7m} the max Doppler reflects the information in Figure \ref{fig:rx_position} and the overall envelope is afected. The envelope in Figure \ref{fig:rx_position_0m} is much noiser than the envelope of Figure \ref{fig:rx_position_7m} and the envelope of this second figure has a more distinct shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore the receiver will be offset from the center of rotation by 7m for the subsequent simulations.

@@

From Figure \ref{fig:pitch_15_135deg} the effect of pitch on the maximum Doppler frequency is a linear increase between a pitch from 0 \textdegree \space to 15\textdegree. Due to the linear increase the pitch of the rotor blade can be accounted for when making measurements on subsequent simulation analysis. One issue does arise in when the transmitter is located underneath the helicopter which is shown in Figure \ref{fig:pitch_tx0}.

@@

The pitch increases linearly at first then climbs at a much faster rate. This offset when the transmitter is directly underneath is caused by the pitch of the blade reflecting rays into the receiver which would be reflected back at the ground if the pitch was 0\textdegree. If the assumption is that the helicopter is hovering when these measurements take place the pitch of the rotor will be much less than the maximum pitch available. This assumption will adjust the maximum pitch to 15\textdegree \space which will allow for the linear assumption to be fulfilled even when the transmitter is directly underneath the helicopter.

@@

In Figure \ref{fig:pitch_tx_elevation_angle} the max Doppler vs elevation angles are initially shifted up in frequency from a pitch of 0\textdegree \space and has changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch envelopes actually become closer to the 0\textdegree \space envelope as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

Where $\theta_{Az}$ is the azimuth angle. The azimuth angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the direction of the rotor blade rotation as it approaches both the receiver and the transmitter and how the rays reflect off the blade as it moves. From the model we see the amount of Doppler form a sinusoidal pattern as the azimuth angle goes from $0$ to $2\pi$. This is mirrored in the simulation shown in Figure \ref{fig:tx_azimuth_angle_200}.

@@

The graph in Figure \ref{fig:tx_azimuth_angle_200} shows both the max and the minimum measured Doppler as the azimuth angle completes a full rotation. The Doppler frequency follows the azimuth angle in a sinusoidal fashion. The sinusoidal pattern is held as the elevation angle changes shown in Figures \ref{fig:tx_azimuth_angle_50} and \ref{fig:tx_azimuth_angle_400}.

@@

%description
The sinusoidal pattern of min and max Doppler frequencies are only present when the receiver is located away from the axis of rotation. This is shown in Figure \ref{fig:tx_azimuth_rx0} which is devoid of any sinusoidal pattern and remains constant as the transmitter azimuth angle is changed.

@@

Pitching the blade modifies the results of the azimuth sweep when compared to Figure \ref{fig:tx_azimuth_angle_200}. The resulting min and max envelope frequencies vs azimuth angle for pitch are shown in \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections differently according to transmitter azimuth. When the transmitter is located between 90\textdegree \space and 270\textdegree \space pitch causes an increase in Doppler but elsewhere decreases the amount of Doppler shift. The decrease is caused by the blade surface top reflecting receding Doppler away from the receiver at those angles.

@@

The effect of azimuth angle on the maximum measured Doppler is characterized in section \ref{sec:transmitter_azimuth_angle_estimation}

@@

The elevation angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver there is no vector component in the direction of rotor travel so the reflected signal is not Doppler shifted. From the model we see the amount of Doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in Figure \ref{fig:tx_elevation_135deg} where the maximum measured Doppler as the elevation angle increases.

>>>>

The rotor blade length from Figure \ref{fig:rx_position_image} is 7.5 m so the receiver will be varied from 0 m to 9 m in 1 m increments. The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of the position of the receiver.

@@

Figure \ref{fig:rx_position} shows maximum Doppler versus the position of the receiver. From this graph, the maximum Doppler frequency changes significantly as the receiver is moved from directly underneath the rotor to past the tip of the rotor at 9 m.The position of the receiver also effects the resulting signal envelope shapes shown in Figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

@@

Comparing Figure \ref{fig:rx_position_0m} with \ref{fig:rx_position_7m} the max Doppler reflects the information in Figure \ref{fig:rx_position} and the overall envelope is affected. The envelope in Figure \ref{fig:rx_position_0m} is much noisier than the envelope of Figure \ref{fig:rx_position_7m} and the envelope of this second figure has a more distinct shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore, the receiver will be offset from the center of rotation by 7 m for the subsequent simulations.

@@

From Figure \ref{fig:pitch_15_135deg} the effect of pitch on the maximum Doppler frequency is a linear increase between a pitch from 0\textdegree \space to 15\textdegree. Due to the linear increase, the pitch of the rotor blade can be accounted for when making measurements on subsequent simulation analysis. One issue does arise in when the transmitter is located underneath the helicopter which is shown in Figure \ref{fig:pitch_tx0}.

@@

The pitch increases linearly at first then climbs at a much faster rate. This offset is caused by the pitch of the blade reflecting rays into the receiver which would be reflected back at the ground if the pitch was 0\textdegree. If the assumption is that the helicopter is hovering when these measurements take place, the pitch of the rotor will be much less than the maximum pitch available. This assumption will adjust the maximum pitch to 15\textdegree \space which will allow for the linear assumption to be fulfilled even when the transmitter is directly underneath the helicopter.

@@

In Figure \ref{fig:pitch_tx_elevation_angle} the max Doppler versus elevation angles are initially shifted up in frequency from a pitch of 0\textdegree \space and, consequently, have changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch envelopes actually become closer to the 0\textdegree \space envelope as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

Where $\theta_{Az}$ is the azimuth angle. The azimuth angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the direction of the rotor blade rotation as it approaches both the receiver and the transmitter and how the rays reflect off the blade as it moves. From the model we see the amount of Doppler form a sinusoidal pattern as the azimuth angle goes from 0\textdegree \space to 360\textdegree. This is mirrored in the simulation shown in Figure \ref{fig:tx_azimuth_angle_200}.

@@

The graph in Figure \ref{fig:tx_azimuth_angle_200} shows both the maximum and the minimum measured Doppler as the azimuth angle completes a full rotation. The Doppler frequency follows the azimuth angle in a sinusoidal fashion. The sinusoidal pattern is held as the elevation angle changes shown in Figures \ref{fig:tx_azimuth_angle_50} and \ref{fig:tx_azimuth_angle_400}.

@@

%description
The sinusoidal pattern of minimum and maximum Doppler frequencies are only present when the receiver is located away from the axis of rotation. This is shown in Figure \ref{fig:tx_azimuth_rx0} which is devoid of any sinusoidal pattern and remains constant as the transmitter azimuth angle is changed.

@@

Pitching the blade modifies the results of the azimuth sweep when compared to Figure \ref{fig:tx_azimuth_angle_200}. The resulting minimum and maximum envelope frequencies versus azimuth angle for pitch are shown in Figure \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections differently according to transmitter azimuth. When the transmitter is located between 90\textdegree \space and 270\textdegree \space pitch causes an increase in Doppler but elsewhere it decreases the amount of Doppler shift. The decrease is caused by the blade surface top reflecting receding Doppler away from the receiver at those angles.

@@

The effect of the azimuth angle on the maximum measured Doppler is characterized in section \ref{sec:transmitter_azimuth_angle_estimation}

@@

The elevation angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver, there is no vector component in the direction of rotor travel so the reflected signal is not Doppler shifted. From the model we see the amount of Doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in Figure \ref{fig:tx_elevation_135deg} where the maximum measured Doppler as the elevation angle increases.
